---
title: "{adamb} Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{adamb-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This package is designed by Adam B, for Adam B. The idea is to package all frequently used
code into a single source for consistency and to make Adam B's life easier.

The package contains fun data sets, useful data visualization functions, and a simplified
way to access personal databases. 

# Getting Started

To get started with the package, install using `remotes`:

```{r}
# install.packages("remotes")

remotes::install_github("adambushman/adamb")
```

Once installed, load the package using the `library` command below or reference
individual components using the `::` framework:

```{r}
# Load the entire library into the session

library("adamb")

# Reference library components

adamb::office_data
```

You'll find helpful functions for data visualization, connecting to databases, 
and even fun, useful data sets used frequently.


# Data Visualization

## Working with Plots

Visualizing data in R via {ggplot2} is a joy but often styling a plot follows
a common convention. With this in mind, use the `theme_adamb()` to leverage
foundational styling via a custom ggplot theme.

```{r}
library('ggplot2')

# Using a plot with no custom theme

ggplot(office_data) +
  geom_point(aes(month, sales))

# Using `theme_adamb()`

ggplot(office_data) +
  geom_point(aes(month, sales)) +
  theme_adamb()
```

For high resolution plots designed for export, use the `start_camcorder()` function
to begin a recording. It's a simplified version of `camcorder::gg_record()` which 
assumes some default values for "device" (.png), "unit" (cm), and "dpi" (300).

```{r}
start_camcorder("/my_directory", 16, 9)
```

This function is designed for use with `theme_adamb_cam()`, which includes sizing
of theme elements appropriate for a recording. A traditional workflow likely involves
using `theme_adamb()` during initial design of the geom or stat layers and then
calling `start_camcorder()` and swapping the theme for `theme_adamb_cam()` to 
finish the look and feel of the plot.

Another nice feature about these custom themes is the ability to pass what we call 
a "color object". Essentially it's a list of named elements containing hexadecimal 
color codes. Here's an example:

```{r}
# A basic color object

color_obj <- list(
  core = "#ac162c", 
  background = "#F5F5F5", 
  text = "#222222", 
  lines = "#F0F0F0"
)

color_obj
```

The theme will reference these named elements to create a great look and feel.
Just pass it to the theme. You can even reference the object's element `core`
to use in the geom or stat layers.

```{r}
# Passing a color object

ggplot(office_data) +
  geom_point(
    aes(month, sales), 
    color = color_obj$core
  ) +
  labs(title = "Office Sales") +
  theme_adamb(color_obj)
```

## Creating Color Objects

You may or may not have a color object in mind for your plot. Consider using our built
in function for creating a color object that is balanced in hue and complimentary in 
elements. Here's how:

Just pass in the type ("Dark" or "Light") and either a hue (0 to 360 representing the color 
wheel) or a valid color name.

```{r}
# Passing in a color hue
gen_color_obj1 = make_color_obj("Dark", hue = 88)

# Passing in a color name
gen_color_obj2 = make_color_obj("Dark", name = "Red")
```

To find all the valid color names, just reference `common_colors` from the package for a
dataframe including their hues:

```{r}
common_colors
```
