---
title: "{adamb} Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{adamb-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This package is designed by Adam B, for Adam B. The idea is to package all frequently used
code into a single source for consistency and to make Adam B's life easier.

The package contains fun data sets, useful data visualization functions, and a simplified
way to access personal databases. 

# Getting Started

To get started with the package, install using `remotes`:

```{r eval=FALSE}
install.packages("remotes")

remotes::install_github("adambushman/adamb", dependencies = TRUE, build_vignettes = TRUE)
```

Once installed, load the package using the `library` command below or reference
individual components using the `::` framework:

```{r}
# Load the entire library into the session

library("adamb")

# Reference library components

adamb::office_data
```

You'll find helpful functions for data visualization, connecting to databases, 
and even fun, useful data sets used frequently.


# Data Visualization

## Working with Plots

Visualizing data in R via {ggplot2} is a joy but often styling a plot follows
a common convention. With this in mind, use the `theme_adamb()` to leverage
foundational styling via a custom ggplot theme.

```{r}
library('ggplot2')

# Using a plot with no custom theme

ggplot(office_data) +
  geom_point(aes(month, sales))

# Using `theme_adamb()`

ggplot(office_data) +
  geom_point(aes(month, sales)) +
  theme_adamb()
```

For high resolution plots designed for export, use the `start_camcorder()` function
to begin a recording. It's a simplified version of `camcorder::gg_record()` which 
assumes some default values for "device" (.png), "unit" (cm), and "dpi" (300).

```{r eval=FALSE}
start_camcorder("/my_directory", 16, 9)
```

This function is designed for use with `theme_adamb_cam()`, which includes sizing
of theme elements appropriate for a recording. A traditional workflow likely involves
using `theme_adamb()` during initial design of the geom or stat layers and then
calling `start_camcorder()` and swapping the theme for `theme_adamb_cam()` to 
finish the look and feel of the plot.

Another nice feature about these custom themes is the ability to pass what we call 
a "color object". Essentially it's a list of named elements containing hexadecimal 
color codes. Here's an example:

```{r}
# A basic color object

color_obj <- list(
  core = "#ac162c", 
  background = "#F5F5F5", 
  text = "#222222", 
  lines = "#F0F0F0"
)

color_obj
```

The theme will reference these named elements to create a great look and feel.
Just pass it to the theme. You can even reference the object's element `core`
to use in the geom or stat layers.

```{r}
# Passing a color object

ggplot(office_data) +
  geom_point(
    aes(month, sales), 
    color = color_obj$core
  ) +
  labs(title = "Office Sales") +
  theme_adamb(color_obj)
```

## Creating Color Objects

You may or may not have a color object in mind for your plot. Consider using our built
in function for creating a color object that is balanced in hue and complimentary in 
elements. Here's how:

Just pass in the type ("Dark" or "Light") and either a hue (0 to 360 representing the color 
wheel) or a valid color name.

```{r}
# Passing in a color hue
gen_color_obj1 = make_color_obj("Dark", hue = 88)
gen_color_obj1

# Passing in a color name
gen_color_obj2 = make_color_obj("Dark", name = "Red")
gen_color_obj2
```

To find all the valid color names, just reference `common_colors` from the package for a
dataframe including their hues:

```{r}
common_colors
```

We can pass one of the above objects to `theme_adamb()` and instantly have a cohesively
styled theme for our plot:

```{r}
ggplot(office_data) +
  geom_point(
    aes(month, sales), 
    color = gen_color_obj2$core
  ) +
  labs(title = "Office Sales") +
  theme_adamb(gen_color_obj2)
```

As per usual, immediately following a custom theme you can add the standard `theme()`
from {ggplot2} to override or add custom elements to the plot.

One of the helper functions to `make_color_obj()` is `hsl_to_hex()`, a function to 
translate "Hue", "Saturation", and "Lightness" values to their hexadecimal representation.
Should this ever be needed, simply call and pass in its 3 values (respecting that "h" must 
be a value between 0 and 360, while "s" and "l" are values between 0 and 100).

```{r}
hsl_to_hex(50, 70, 90)
```

Often in working with visuals, we have a background color that may not provide enough contrast for readability. In such a case, we can use the `contrast_text_color()` function to get a better color.

Simply pass a hexadecimal code to the function and it will determine if white (#FFFFFF) or black (#000000) is the best contrasting color and returns the code.

```{r}
contrast_text_color("#AC192D")
```


# Querying Databases

Databases often contain the data for our analyses and visualizations. Extract that data
is one of the core first steps. However, most packages have a pretty heavy workflow for
connection and any operations. 

This package aims to simplify it quite a bit by encapsulating a good portion of the 
workflow and using reasonable defaults that can be overridden if possible. It is noted these
functions are designed solely for use with MySQL databases.

## Connecting

The headliner here is `mysql_connect()`, which handles all the leg work of connection with
some common defaults:

```{r eval=FALSE}
# Pass the minimum set of parameters
mysql_connect(username = "my_user", password = "my_pass123", dbname = "my_db")

# Pass all parameters to override the defaults
mysql_connect(
  username = "my_user", 
  password = "my_pass123", 
  dbname = "my_db", 
  host = "my_host", 
  port = 12345, 
  cname = "my_connection"
)
```

Using the minimum required parameters will leverage default "host" and "port" values 
for Digital Ocean (my main database server).

"cname" stands for the connection variable name to be assigned to the environment. 
If no value is passed, it will default to "con". For functions below, use whatever name
for the connection appears in the environment after `mysql_connect()`.

## Peeking

Two functions help describe the database before diving head first. Using whatever connection
variable from the environment created by `mysql_connect()` (let's say "con"), we can 
get a summary of tables and a singular table's fields:

```{r eval=FALSE}
# Peek at all tables in the database
mysql_peek_tables(con)

# Peek at all the fields in a particular table
mysql_peek_fields(con, "table_1")
```

With such detail in hand, we can begin to execute query's on the database.

## Querying

To send a query, we need only reference the connection variable and pass a string
containing our query.

```{r eval=FALSE}
my_query = "SELECT * FROM table_1"

mysql_send_query(con, my_query)
```

This function encapsulates all the logic for requesting a response, storing the data, and
clearing the response. We can move right along with our analysis.

## Finishing Up

Once we've completed all of our operations on the database, we need to close the connection. 
While the normal workflow with the direct API is simple enough, you can use the built in 
function here:

```{r eval=FALSE}
mysql_close(con)
```

# Data

Two data sets have been included in this package for complimentary use with the functions
therein:

*   `common_colors`
    -   This is a dataset containing the common color names and hue codes for the color wheel
    at 30 deg intervals. It's commonly used with `make_color_obj()`.
*   `office_sales`
    -   This is a nice dataset for practicing various chart types in {ggplot2}. It's derived 
    from *The Office* TV show, S07:E14, which displays the monthly sales figures for office and
    warehouse staff.

# Coming Soon

More features of the package will debut soon!
